---
title: "eds213_data_cleaning_assign_Marie"
format: html
---

# Load in Libraries
```{r}
#| message: false

# Libraries
library(tidyverse)

# file names
datadir_raw <- file.path("data","raw")

datadir_processed <- file.path("data", "processed/")

# Read in Data
starting_snowsurvey_csv <- read_csv(here::here("data", "processed", "snow_cover.csv"))

# Quick view of data
glimpse(starting_snowsurvey_csv)

```
# View values in the Water Column
```{r}
# Non numeric values
starting_snowsurvey_csv |>
  count(Water_cover) |>
  filter(is.na(as.numeric(Water_cover)))
```
---
# Water Cover
Remove non-numeric values and turn them into NA
```{r}
water_fixed <- starting_snowsurvey_csv |>
  mutate(Water_cover = ifelse(Water_cover == "-", NA, Water_cover),
         Water_cover = ifelse(Water_cover == ".", NA, Water_cover),
         Water_cover = ifelse(Water_cover == "n/a", NA, Water_cover),
         Water_cover = ifelse(Water_cover == "unk", NA, Water_cover))

water_fixed|>
  count(Water_cover) |>
  filter(is.na(as.numeric(Water_cover)))

class(water_fixed$Water_cover)
```


Convert to Numeric
```{r}
# Convert Column from Character to  Numeric
water_fixed <- water_fixed |>
  mutate(Water_cover = as.numeric(Water_cover),
         # Make NA values for greater than 100
          Water_cover = ifelse(Water_cover > 100, NA, Water_cover))
# View class 
class(water_fixed$Water_cover)

# Quick Summary Stats
summary(water_fixed$Water_cover)

# View Data
glimpse(water_fixed)
```
---
# Land Cover
View non-numeric values within `land_cover` column
```{r}
water_fixed |>
  count(Land_cover) |>
  filter(is.na(as.numeric(Land_cover)))
```

# Remove Non-numeric Values from Land_cover
```{r}
land_fixed <- water_fixed |>
  # Remove values that are not numeric
  mutate(Land_cover = ifelse(Land_cover == "-", NA, Land_cover),
         Land_cover = ifelse(Land_cover == ".", NA, Land_cover),
         Land_cover = ifelse(Land_cover == "n/a", NA, Land_cover),
         Land_cover = ifelse(Land_cover == "unk", NA, Land_cover))

# View NA values
land_fixed|>
  count(Land_cover) |>
  filter(is.na(as.numeric(Land_cover)))

# View unique values in column
unique(land_fixed$Land_cover)



```

# Convert Column to Numeric
```{r}
land_fixed <- land_fixed %>% 
  # Switch Character to Numeric
  mutate(Land_cover = as.numeric(Land_cover)) %>% 
  # Remove Negative values
  mutate(Land_cover = ifelse(Land_cover %in% c(-100, -298), NA, Land_cover)) %>% 
  mutate(Land_cover = ifelse(Land_cover == "NA", NA, Land_cover),
         Land_cover = ifelse(Land_cover > 100, NA, Land_cover))



# Confirm column is now numeric
class(land_fixed$Land_cover)

# View unique values in column
unique(land_fixed$Land_cover)

```


---
# Total Cover 
View Non-numeric values in `Total_cover` Column
```{r}
land_fixed |>
  count(Total_cover) |>
  filter(is.na(as.numeric(Total_cover)))
```

# Remove Non-numeric values and Get a count of NA values 
```{r}
total_fixed <- land_fixed |>
  # Remove values that are not numeric
  mutate(Total_cover = ifelse(Total_cover == "-", NA, Total_cover),
         Total_cover = ifelse(Total_cover == ".", NA, Total_cover),
         Total_cover = ifelse(Total_cover == "n/a", NA, Total_cover),
         Total_cover = ifelse(Total_cover == "<row r=\\64\\ spans=\\1:11\\ x14ac:dyDescent=\\0.2\\>", NA, Total_cover))

# Retrieving NA count
total_fixed|>
  count(Total_cover) |>
  filter(is.na(as.numeric(Total_cover)))


```

# Convert to Numeric, Verify Updated Class, and Check Unique Values
```{r}
total_fixed <- total_fixed %>% 
  # Switch Character to Numeric
  mutate(Total_cover = as.numeric(Total_cover)) %>% 
  # Remove values over 100
  mutate(Total_cover = ifelse(Total_cover > 100, NA, Total_cover))

# Confirm column is now numeric
class(total_fixed$Total_cover)

# Let's get an understanding of values in this column
unique(total_fixed$Total_cover)

```

# Export Finding into CSV
```{r}
# check if the folder exists
dir.create(datadir_processed, showWarnings = FALSE)

# write the file
write_csv(total_fixed, file.path(datadir_processed, "all_cover_fixed_MARIE.csv"))
```




