---
title: "eds213_data_cleaning_assign_Marie"
format: html
---

# Load in Libraries
Adding in code chunk from class to have the same file path for data
```{r}
#| message: false

# Libraries
library(tidyverse)

# file names
datadir_raw <- file.path("data","raw")

datadir_processed <- file.path("data", "processed/")

# Read in Data- using here::here()
starting_snowsurvey_csv <- read_csv(here::here("data", "processed", "snow_cover.csv"))

# Quick view of data
glimpse(starting_snowsurvey_csv)
```
# View values in the Water Column
```{r}
# View Non-numeric values
starting_snowsurvey_csv |>
  count(Water_cover) |>
  filter(is.na(as.numeric(Water_cover)))
```
---
# Water Cover
Assign into new dataframe. Replace `-`, `.`, `unk`, and `n/a` are non-numeric placeholders for missing data with NA
```{r}
water_fixed <- starting_snowsurvey_csv |>
  mutate(Water_cover = ifelse(Water_cover == "-", NA, Water_cover),
         Water_cover = ifelse(Water_cover == ".", NA, Water_cover),
         Water_cover = ifelse(Water_cover == "n/a", NA, Water_cover),
         Water_cover = ifelse(Water_cover == "unk", NA, Water_cover))

# View total count of NA after removing non-numeric values
water_fixed|>
  count(Water_cover) |>
  filter(is.na(as.numeric(Water_cover)))

# View datatype
class(water_fixed$Water_cover)
```


# Convert `Water_cover` from character to numeric.
Values greater than 100 are invalid for a percentage and are replaced with NA.
```{r}
# Convert Column from Character to  Numeric
water_fixed <- water_fixed |>
  mutate(Water_cover = as.numeric(Water_cover),
         # Make NA values for greater than 100
          Water_cover = ifelse(Water_cover > 100, NA, Water_cover))
# View class 
class(water_fixed$Water_cover)

# Quick Summary Stats- View Max and Min Value
summary(water_fixed$Water_cover)

# View Data
glimpse(water_fixed)
```

---
# Land Cover
```{r}
# View Non-numeric values
water_fixed |>
  count(Land_cover) |>
  filter(is.na(as.numeric(Land_cover)))
```

# Remove Non-numeric Values from Land_cover
Assign into new dataframe. Replace `-`, `.`, `unk`, and `n/a` are non-numeric placeholders for missing data with NA
```{r}
land_fixed <- water_fixed |>
  # Remove values that are not numeric
  mutate(Land_cover = ifelse(Land_cover == "-", NA, Land_cover),
         Land_cover = ifelse(Land_cover == ".", NA, Land_cover),
         Land_cover = ifelse(Land_cover == "n/a", NA, Land_cover),
         Land_cover = ifelse(Land_cover == "unk", NA, Land_cover))

# View NA values
land_fixed|>
  count(Land_cover) |>
  filter(is.na(as.numeric(Land_cover)))

# View unique values in column
unique(land_fixed$Land_cover)



```

# Convert `Land_cover` from character to numeric.
Values greater than 100 or less than zero are invalid for a percentage and are replaced with NA.
```{r}
land_fixed <- land_fixed %>% 
  # Switch Character to Numeric
  mutate(Land_cover = as.numeric(Land_cover)) %>% 
  # Remove Negative values
  mutate(Land_cover = ifelse(Land_cover < 0, NA, Land_cover),
  # Remove NA values
  Land_cover = ifelse(Land_cover == "NA", NA, Land_cover),
  # Remove any value over 100
  Land_cover = ifelse(Land_cover > 100, NA, Land_cover))


# Confirm column is now numeric
class(land_fixed$Land_cover)

# View unique values in column
unique(land_fixed$Land_cover)

```


---
# Total Cover 
```{r}
land_fixed |>
  count(Total_cover) |>
  filter(is.na(as.numeric(Total_cover)))
```

# Convert `Total_cover` from character to numeric.
Assign into new dataframe. Replace `-`, `.`, `n/a`, and long string are non-numeric placeholders for missing data with NA
```{r}
total_fixed <- land_fixed |>
  # Remove values that are not numeric
  mutate(Total_cover = ifelse(Total_cover == "-", NA, Total_cover),
         Total_cover = ifelse(Total_cover == ".", NA, Total_cover),
         Total_cover = ifelse(Total_cover == "n/a", NA, Total_cover),
         Total_cover = ifelse(Total_cover == "<row r=\\64\\ spans=\\1:11\\ x14ac:dyDescent=\\0.2\\>", NA, Total_cover))

# Retrieving NA count
total_fixed|>
  count(Total_cover) |>
  filter(is.na(as.numeric(Total_cover)))


```

# Convert to Numeric, Verify Updated Class, and Check Unique Values
Rows that did not add up to 100 or had NA values were removed from the dataframe. I decided to remove these rows since it would be inconsistent with the other rows and also to avoid any confusion. This is a more clear approach with a reassurance that all the values in each column of Snow, Water, and Land add up to a Total cover of 100. A total of 14 rows were removed where snow, water, and land cover did not sum to 100. Two patterns were found; rows where all three values were 0, but total_cover was 100, and rows where the three columns summed to 90.
```{r}
# Let's try another attempt
total_fixed2<- total_fixed %>%
  # Switch Character to Numeric
  mutate(Total_cover = as.numeric(Total_cover)) %>%
  # Replace any non-100 total with NA
  mutate(Total_cover = ifelse(Total_cover != 100, NA, Total_cover)) %>%
  # If Total_cover is missing but all three cover columns are known, compute it as their sum
  mutate(Total_cover = ifelse(
    is.na(Total_cover) & !is.na(Snow_cover) & !is.na(Water_cover) & !is.na(Land_cover),
    Snow_cover + Water_cover + Land_cover,
    Total_cover)) %>%
  # Only keep Total_cover if it equals 100
  mutate(Total_cover = ifelse(Total_cover == 100, 100, NA)) %>% 
  # Remove rows that do not add up to 100 
  filter(Snow_cover + Water_cover + Land_cover == 100)

# Verify column is now numeric
class(total_fixed2$Total_cover)

```

# Do all column covers add up to 100 in Total Cover?
```{r}
# Check for any NA values in the four cover columns
sum(is.na(total_fixed2[, c("Snow_cover", "Water_cover", "Land_cover", "Total_cover")]))

# Check all rows sum to 100
print(paste("All rows sum to 100:", all(total_fixed2$Snow_cover + total_fixed2$Water_cover + total_fixed2$Land_cover == 100)))
```


# Export Finding into CSV
```{r}
# check if the folder exists
dir.create(datadir_processed, showWarnings = FALSE)

# write the file
write_csv(total_fixed2, file.path(datadir_processed, "all_cover_fixed_MARIE.csv"))
```




